\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-20 forwards-and-backwards.tex
generator convergence appendix

Putting these together,
the level $u$ of an individual at $x$ in $\lp$ evolves according to:
$$
    \dot u
    =
    - \frac{\theta}{\lambda} u \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \theta \mu(x,\eta) u .
$$
The factor of $\theta$ appears here because we have rescaled time.
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
Recall from equation \eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$,
and so we will have that
\begin{align*}
b(x, \eta)
&=
    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) q(x, dy)
    + F(x, \eta) \\
&\to
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) \qquad \text{as } \theta \to \infty .
\end{align*}

@@

\comment{Here's an informal writeup of this convergence;
mostly I'm doing it to make sure I've got the right stuff above.}

@@

First, the contribution from births is
\begin{align*}
\begin{split}
f(\lp)
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 \lambda}
        \int_u^\lambda
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{\lambda}
        \int_u^\lambda \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, u_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align*}
We have that
\begin{align*}
    \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
    \to
    \DG(gr)(x) \qquad \text{as }\theta \to \infty,
\end{align*}
i.e., to $\DG\left(g(\cdot, u) r(\cdot, \smooth{r}\eta(\cdot))\right)(x)$.
Furthermore, by our assumption on $g$,
$\int_u^\lambda g(x, u_1) du_1 / \lambda \to 1$ for all $x$ and $u$,
so the first term here converges to
\begin{align*}
    2 \gamma(x, \eta)
        \frac{1}{2}
        \frac{
            \DG\left( g(\cdot, u) r(\cdot, \smooth{r}\eta(\cdot))\right)(x)
        }{
            g(x, u)
        } .
\end{align*}
This corresponds to each particle moving according to the motion induced by
$g \mapsto \gamma \DG(gr)$.
\comment{TODO: explain the $r$-tilted $\DG$ motion somewhere.}
As for the second term,
since $\int g(y, u_1) q(x, dy) \to g(x, u_1)$
and $\theta/\lambda \to \delta$,
it converges to
\begin{align*}
    \delta
    \int_u^\infty
    \left( g(x, u_1) - 1 \right)
    r(x, \eta) .
\end{align*}
This corresponds to births at higher levels at rate $\gamma r$.

>>>>

Putting these together,
the level $u$ of an individual at $x$ in $\lp$ evolves according to:
$$
    \dot u
    =
    - \frac{\theta}{\lambda} u \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \theta \mu(x,\eta) u .
$$
The factor of $\theta$ appears here because we have rescaled time.
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
Recall from equation \eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$,
and so we will have that
\begin{align} \label{eqn:b_limit}
b(x, \eta)
&=
    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) q(x, dy)
    + F(x, \eta) \\
&\to
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) \qquad \text{as } \theta \to \infty .
\end{align}

@@

\comment{
    I did this informal writeup of convergence of the lookdown generator
    without looking at the previous version
    mostly to double-check what we have there
    (which is probably better).
}

@@

First, the contribution from births is
\begin{align*}
\begin{split}
f(\lp)
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 \lambda}
        \int_u^\lambda
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{\lambda}
        \int_u^\lambda \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, u_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align*}
We have that
\begin{align*}
    \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
    \to
    \DG(gr)(x) \qquad \text{as }\theta \to \infty,
\end{align*}
i.e., to $\DG\left(g(\cdot, u) r(\cdot, \smooth{r}\eta(\cdot))\right)(x)$.
Furthermore, by our assumption on $g$,
$\int_u^\lambda g(x, u_1) du_1 / \lambda \to 1$ for all $x$ and $u$,
so the first term here converges to
\begin{align*}
    \gamma(x, \eta)
        \frac{
            \DG\left( g(\cdot, u) r(\cdot, \smooth{r}\eta(\cdot))\right)(x)
        }{
            g(x, u)
        } .
\end{align*}
This corresponds to each particle moving according to the motion induced by
$g \mapsto \gamma \DG(gr)$.
\comment{TODO: explain the $r$-tilted $\DG$ motion somewhere.}
As for the second term,
since $\int g(y, u_1) q(x, dy) \to g(x, u_1)$
and $\theta/\lambda \to \delta$,
it converges to
\begin{align*}
    2 \delta
    \gamma(x, \eta)
    r(x, \eta) 
    \int_u^\infty
    \left( g(x, u_1) - 1 \right)
    du_1
\end{align*}
This corresponds to births at higher levels at rate $2 \delta \gamma r$.

The remaining term is \eqref{eqn:level_generator}:
\begin{align*}
    f(\lp)
    \mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    \left(
    \theta
        \lambda^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        b(x, \eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align*}
The first part only uses the fact that $\theta/\lambda \to \delta$,
while as noted in \eqref{eqn:b_limit}, $b$ converges to $\gamma \DG r + F$,
so in total this term converges to
\begin{align*}
    f(\lp)
    \mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    \left(
    \delta
        \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        \left\{
            \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta)
        \right\} u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align*}
In other words, the level of a particle at $x$ evolves according to
\begin{align*}
    \dot u
    =
    \delta
        \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        \left\{
            \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta)
        \right\} u .
\end{align*}
